### 2018-01-23 submission/manuscript.Rmd
change the title
update figure legends

# Investigating the Microbiota and Colorectal Cancer: The Importance of Community

@@

**Figure 3: CRC-Associated Genera Relative Risk for Adenoma and Carcinoma in Stool and Tissue.** A) Adenoma relative risk in stool. B) Carcinoma relative risk in stool. C) Adenoma relative risk in tissue. D) Carcinoma relative risk in tissue. For all panels the relative risk was also compared to whether one, two, three, or four of the CRC-associated genera were present.

**Figure 4: OTU Random Forest Model of Stool Across Studies.** A) Adenoma random forest model between the full community and CRC-associated genera OTUs only. B) Carcinoma random forest model between the full community and CRC-associated genera OTUs only. The dotted line represents an AUC of 0.5 and the lines represent the range in which the AUC for the 100 different 80/20 runs fell between. The solid red line represents the median AUC of all the studies for either the full community or CRC-associated genera OTUS only model.

**Figure 5: OTU Random Forest Model of Tissue Across Studies.** A) Adenoma random forest model between the full community and CRC-associated genera OTUs only. B) Carcinoma random forest model between the full community and CRC-associated genera OTUs only. The dotted line represents an AUC of 0.5 and the lines represent the range in which the AUC for the 100 different 80/20 runs fell between. The solid red line represents the median AUC of all the studies for either the full community or CRC-associated genera OTUS only model.

@@

**Figure S2: Random Forest Genus Model AUC for each Stool Study.** A) AUC of adenoma models using all genera or CRC-associated genera only. B) AUC of carcinoma models using all genera or CRC-associated genera only. The black line represents the median within each group.

**Figure S3: Random Forest Genus Model AUC for each Tissue Study.** A) AUC of adenoma models using all genera or only CRC-associated genera divided between matched and unmatched tissue. B) AUC of carcinoma models using all genera or CRC-associated genera only. The black line represents the median within each group divided between matched and unmatched tissue.

**Figure S4: Random Forest Prediction Success Using Genera for each Stool Study.** A) AUC for prediction in adenoma using all genera or CRC associated genera only. B) AUC for prediction in carcinoma using all genera or CRC-associated genera only. The dotted line represents an AUC of 0.5. The x-axis is the data set in which the model was initially trained on. The red lines represent the median AUC using that specific study as the training set.

**Figure S5: Random Forest Prediction Success of Carcinoma Using Genera for each Tissue Study.** A) AUC for prediction in unmatched tissue for all genera or CRC-associated genera only. B) AUC for prediction in matched tissue using all genera or CRC-associated genera only. The dotted line represents an AUC of 0.5. The x-axis is the data set in which the model was initially trained on. The red lines represent the median AUC using that specific study as the training set.

**Figure S6: Random Forest Prediction Success of Adenoma Using Genera for each Tissue Study.** The red lines represent the median AUC using that specific study as the training set.

>>>>

# Making Sense of the Noise: Leveraging Existing 16S rRNA Gene Surveys to Identify Key Community Members in Colorectal Cancer

@@

**Figure 3: Top 5 Genera that Decrease and Increase Relative Risk for Lesion.** A) Adenoma relative risk in stool. B) Carcinoma relative risk in stool. C) Adenoma relative risk in tissue. D) Carcinoma relative risk in tissue. For all panels the relative risk was also compared to whether one, two, three, or four of the CRC-associated genera were present.

**Figure 4: Stool OTU Random Forest Model Across Studies.** A) Adenoma random forest model between the full community and select genera OTUs only. B) Carcinoma random forest model between the full community and select genera OTUs only. The dotted line represents an AUC of 0.5 and the lines represent the range in which the AUC for the 100 different 80/20 runs fell between. The solid red line represents the median AUC of all the studies for either the full community or select genera OTUS only model.

**Figure 5: Tissue OTU Random Forest Model Across Studies.** A) Adenoma random forest model between the full community and select genera OTUs only. B) Carcinoma random forest model between the full community and select genera OTUs only. The dotted line represents an AUC of 0.5 and the lines represent the range in which the AUC for the 100 different 80/20 runs fell between. The solid red line represents the median AUC of all the studies for either the full community or select genera OTUS only model.

@@

**Figure S2: Stool Random Forest Genus Model AUC for each Study.** A) AUC of adenoma models using all genera or select genera only. B) AUC of carcinoma models using all genera or select genera only. The black line represents the median within each group.

**Figure S3: Tissue Random Forest Genus Model AUC for each Study.** A) AUC of adenoma models using all genera or only select genera divided between matched and unmatched tissue. B) AUC of carcinoma models using all genera or select genera only. The black line represents the median within each group divided between matched and unmatched tissue.

**Figure S4: Stool Random Forest Prediction Success Using Genera Across Studies.** A) AUC for prediction in adenoma using all genera or select genera only. B) AUC for prediction in carcinoma using all genera or select genera only. The dotted line represents an AUC of 0.5. The x-axis is the data set in which the model was initially trained on. The red lines represent the median AUC using that specific study as the training set.

**Figure S5: Tissue Random Forest Prediction Success of Carcinoma Using Genera Across Studies.** A) AUC for prediction in unmatched tissue for all genera or select genera only. B) AUC for prediction in matched tissue using all genera or select genera only. The dotted line represents an AUC of 0.5. The x-axis is the data set in which the model was initially trained on. The red lines represent the median AUC using that specific study as the training set.

**Figure S6: Tissue Random Forest Prediction Success of Adenoma Using Genera Across Studies.** The red lines represent the median AUC using that specific study as the training set.

### 2018-01-23 submission/manuscript.tex
update figure legends

\textbf{\emph{Statistical Analysis:}} All statistical analysis after
sequence processing utilized the R (v3.4.3) software package {[}38{]}.
For the \(\alpha\)-diversity analysis, values were power transformed
using the rcompanion (v1.10.1) package {[}39{]} and then Z-score
normalized using the car (v2.1.5) package {[}40{]}. Testing for
\(\alpha\)-diversity differences utilized linear mixed-effect models
created using the lme4 (v1.1.14) package {[}41{]} to correct for study,
repeat sampling of individuals (tissue only), and 16S hypervariable
region used. Relative risk was analyzed using both the epiR (v0.9.87)
and metafor (v2.0.0) packages {[}42,43{]} by assessing how many
individuals with and without disease were above and below the overall
median value within each specific study. Relative risk significance
testing utilized the chi-squared test. \(\beta\)-diversity differences
utilized a Bray-Curtis distance matrix and PERMANOVA executed with the
vegan (v2.4.4) package {[}44{]}. Random Forest models were built using
both the caret (v6.0.77) and randomForest (v4.6.12) packages
{[}45,46{]}. Power analysis and estimations were made using the pwr
(v1.2.1) and statmod (v1.4.30) packages {[}47,48{]}. All figures were
created using both ggplot2 (v2.2.1) and gridExtra (v2.3) packages
{[}49,50{]}.

@@

\textbf{Figure 3: CRC-Associated Genera Relative Risk for Adenoma and
Carcinoma in Stool and Tissue.} A) Adenoma relative risk in stool. B)
Carcinoma relative risk in stool. C) Adenoma relative risk in tissue. D)
Carcinoma relative risk in tissue. For all panels the relative risk was
also compared to whether one, two, three, or four of the CRC-associated
genera were present.

\textbf{Figure 4: OTU Random Forest Model of Stool Across Studies.} A)
Adenoma random forest model between the full community and
CRC-associated genera OTUs only. B) Carcinoma random forest model
between the full community and CRC-associated genera OTUs only. The
dotted line represents an AUC of 0.5 and the lines represent the range
in which the AUC for the 100 different 80/20 runs fell between. The
solid red line represents the median AUC of all the studies for either
the full community or CRC-associated genera OTUS only model.

\textbf{Figure 5: OTU Random Forest Model of Tissue Across Studies.} A)
Adenoma random forest model between the full community and
CRC-associated genera OTUs only. B) Carcinoma random forest model
between the full community and CRC-associated genera OTUs only. The
dotted line represents an AUC of 0.5 and the lines represent the range
in which the AUC for the 100 different 80/20 runs fell between. The
solid red line represents the median AUC of all the studies for either
the full community or CRC-associated genera OTUS only model.

@@

\textbf{Figure S2: Random Forest Genus Model AUC for each Stool Study.}
A) AUC of adenoma models using all genera or CRC-associated genera only.
B) AUC of carcinoma models using all genera or CRC-associated genera
only. The black line represents the median within each group.

\textbf{Figure S3: Random Forest Genus Model AUC for each Tissue Study.}
A) AUC of adenoma models using all genera or only CRC-associated genera
divided between matched and unmatched tissue. B) AUC of carcinoma models
using all genera or CRC-associated genera only. The black line
represents the median within each group divided between matched and
unmatched tissue.

\textbf{Figure S4: Random Forest Prediction Success Using Genera for
each Stool Study.} A) AUC for prediction in adenoma using all genera or
CRC associated genera only. B) AUC for prediction in carcinoma using all
genera or CRC-associated genera only. The dotted line represents an AUC
of 0.5. The x-axis is the data set in which the model was initially
trained on. The red lines represent the median AUC using that specific
study as the training set.

\textbf{Figure S5: Random Forest Prediction Success of Carcinoma Using
Genera for each Tissue Study.} A) AUC for prediction in unmatched tissue
for all genera or CRC-associated genera only. B) AUC for prediction in
matched tissue using all genera or CRC-associated genera only. The
dotted line represents an AUC of 0.5. The x-axis is the data set in
which the model was initially trained on. The red lines represent the
median AUC using that specific study as the training set.

\textbf{Figure S6: Random Forest Prediction Success of Adenoma Using
Genera for each Tissue Study.} The red lines represent the median AUC
using that specific study as the training set.

>>>>

\textbf{\emph{Statistical Analysis:}} All statistical analysis after
sequence processing utilized the R (v3.4.3) software package {[}38{]}.
For the \(\alpha\)-diversity analysis, values were power transformed
using the rcompanion (v1.11.1) package {[}39{]} and then Z-score
normalized using the car (v2.1.6) package {[}40{]}. Testing for
\(\alpha\)-diversity differences utilized linear mixed-effect models
created using the lme4 (v1.1.15) package {[}41{]} to correct for study,
repeat sampling of individuals (tissue only), and 16S hypervariable
region used. Relative risk was analyzed using both the epiR (v0.9.93)
and metafor (v2.0.0) packages {[}42,43{]} by assessing how many
individuals with and without disease were above and below the overall
median value within each specific study. Relative risk significance
testing utilized the chi-squared test. \(\beta\)-diversity differences
utilized a Bray-Curtis distance matrix and PERMANOVA executed with the
vegan (v2.4.5) package {[}44{]}. Random Forest models were built using
both the caret (v6.0.78) and randomForest (v4.6.12) packages
{[}45,46{]}. Power analysis and estimations were made using the pwr
(v1.2.1) and statmod (v1.4.30) packages {[}47,48{]}. All figures were
created using both ggplot2 (v2.2.1) and gridExtra (v2.3) packages
{[}49,50{]}.

@@

\textbf{Figure 3: Top 5 Genera that Decrease and Increase Relative Risk
for Lesion.} A) Adenoma relative risk in stool. B) Carcinoma relative
risk in stool. C) Adenoma relative risk in tissue. D) Carcinoma relative
risk in tissue. For all panels the relative risk was also compared to
whether one, two, three, or four of the CRC-associated genera were
present.

\textbf{Figure 4: Stool OTU Random Forest Model Across Studies.} A)
Adenoma random forest model between the full community and select genera
OTUs only. B) Carcinoma random forest model between the full community
and select genera OTUs only. The dotted line represents an AUC of 0.5
and the lines represent the range in which the AUC for the 100 different
80/20 runs fell between. The solid red line represents the median AUC of
all the studies for either the full community or select genera OTUS only
model.

\textbf{Figure 5: Tissue OTU Random Forest Model Across Studies.} A)
Adenoma random forest model between the full community and select genera
OTUs only. B) Carcinoma random forest model between the full community
and select genera OTUs only. The dotted line represents an AUC of 0.5
and the lines represent the range in which the AUC for the 100 different
80/20 runs fell between. The solid red line represents the median AUC of
all the studies for either the full community or select genera OTUS only
model.

@@

\textbf{Figure S2: Stool Random Forest Genus Model AUC for each Study.}
A) AUC of adenoma models using all genera or select genera only. B) AUC
of carcinoma models using all genera or select genera only. The black
line represents the median within each group.

\textbf{Figure S3: Tissue Random Forest Genus Model AUC for each Study.}
A) AUC of adenoma models using all genera or only select genera divided
between matched and unmatched tissue. B) AUC of carcinoma models using
all genera or select genera only. The black line represents the median
within each group divided between matched and unmatched tissue.

\textbf{Figure S4: Stool Random Forest Prediction Success Using Genera
Across Studies.} A) AUC for prediction in adenoma using all genera or
select genera only. B) AUC for prediction in carcinoma using all genera
or select genera only. The dotted line represents an AUC of 0.5. The
x-axis is the data set in which the model was initially trained on. The
red lines represent the median AUC using that specific study as the
training set.

\textbf{Figure S5: Tissue Random Forest Prediction Success of Carcinoma
Using Genera Across Studies.} A) AUC for prediction in unmatched tissue
for all genera or select genera only. B) AUC for prediction in matched
tissue using all genera or select genera only. The dotted line
represents an AUC of 0.5. The x-axis is the data set in which the model
was initially trained on. The red lines represent the median AUC using
that specific study as the training set.

\textbf{Figure S6: Tissue Random Forest Prediction Success of Adenoma
Using Genera Across Studies.} The red lines represent the median AUC
using that specific study as the training set.
